R Markdown for Reproducible Research

### 2019-01-06 paper.Rmd
add explanatory r markdown section

If Markdown is a lightweight markup language and text formatting syntax, R Markdown
must be Markdown + R code, right? Yes, with a catch. R Markdown actually supports
_many_ programming languages and output types. It got its start with R, and R is
the primary language engine, but R is an interface language. This means that R
is commonly used to "wrap" or

@@

## WHAT ABOUT OTHER LANGUAGES?

## OUTPUT TYPES

@@

>>>>

If Markdown is a lightweight markup language and text formatting syntax, R
Markdown must be Markdown + R code, right? Yes, with a catch. R Markdown
actually supports _many_ programming languages and output types. It got its
start with R, and R is the primary language engine, but R is an interface
language. This means that R has been (from its outset) commonly used to "wrap"
or call other programming languages. This began as Fortran and C, but has
expanded to include C++, Python, JavaScript, SQL, Stan, and more. As a result, R
Markdown supports the inclusion of many of these programming languages.

## THE YAML HEADER

We will focus on R code for now, and come back to the other languages later. R
Markdown has a handful of noteworthy differences from general Markdown. First,
every R Markdown document begins with a YAML header:

```
---
title: "R Markdown for Reproducible Research"
output: pdf_document
---
```

The YAML header uses YAML syntax to declare all sorts of information about the
document. The full documentation on what options are available is [here](todo),
but the primary elements are the document's Title and Output type. The output type
has many options, and can actually support multiple options at the same time.

Some choices for Output Type:

- Static HTML page
- HTML website
- PDF document
- Microsoft Word document
- HTML Presentation
- PDF Presentation
- Microsoft PowerPoint Presentation
- Shiny (interactive) web application
- Blog, Book, etc.

## THE REST OF THE DOCUMENT

Once the YAML header is defined, the body of the document contains:

- Markdown text
- inline code
- code chunks
- embedded items (images, etc.)

Inline code is written like ` r my_function ` and gets placed into the document
like text. It can be very useful for displaying data-determined values
programmatically so that they are dynamically changed when the document is
regenerated (for instance, if new data is acquired).

Code chunks are much like inline code, but they stand alone and are more useful for
lengthier code blocks that do complex logic, setup, value preparation, plotting, or
other output generation.

Embedded images and other resources are included into R Markdown much like they
would be in an HTML page. The syntax looks like `![Text alias](path/to/resource)`.

Once the document is written, it is passed through the `knitr` package to "knit"
the R Markdown document into its final output(s). This creates a new
/ fresh R context to render the document and thereby helps ensure
reproducibility.

@@

## WHAT ABOUT THE OTHER LANGUAGES?

@@

### TODO
- Link for YAML syntax
- Link for YAML header docs
- Resource on R as an interface language
